Fast and optimal centroiding of faint stars

### 2015-02-17 ms/figure.tex
Update figure.tex

In this paper, we study how optimal various techniques are at centroiding 
faint stars. That is, we apply a number of centroiding methods 
to a large number of simulated stars with different signal-to-noise ratio and
 size realizations. Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

Given the analytic model for the PSF,
we derive an expression for the fundamental lower bound on the centroiding error as
a function of the parameters of the PSF model (\eg size),
and signal-to-noise-ratio of stars. We create two sets of simulations for which we 
know the CRLB, one with variable (constant) signal-to-noise ratio (FWHM), and one 
with variable (constant) FWHM (signal-to-noise ratio). After applying
different centroiding methods to the simulations,
we investigate how close these methods can get to saturating the CRLB for
various ranges of background noise and FWHM.

In this work, we focus on three centroiding methods. The first method finds
 the centroid of stars by smoothing the image of
 stars by a Gaussian kernel of a fixed size, and fitting a second order
 polynomial to the $3\times$3 patch around the brightest pixel of smoothed
 image. This method is fast and does not require any knowledge of the 
PSF. In the second method, we apply the same polynomial centroiding to the
 $3\times$3 patch around the brightest pixel of the image of star convolved 
with the full PSF postage stamp. The last method that we consider, is fitting 
a PSF profile, assuming that we have a good PSF estimate, to the star.
%We compare the accuracy of this method with, the naive 3$\times$3 second-order 
%polynomial method, the 3$\times$3 quartic approximation
%method used in SDSS photometric pipeline, and fitting a PSF model to the 
%star.

This paper is structured as follows. In section \ref{sec:method},
we give a brief overview of the centroiding methods used in our investigation.
In section \ref{sec:data}, we discuss the simulated data, and the CRLB on centroiding error.
In section \ref{sec:results}, we compare the results of our proposed method with 
those obtained from centroiding methods discussed in \ref{sec:method}, and 
with the CRLB derived in \ref{sec:data}. Finally, we discuss and conclude
in \ref{sec:discussion}.

\section{Methods}\label{sec:method}

@@

Now, we modify the simple polynomial method explained above,
by smoothing the image of the star with a Gaussian kernel
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}).
\eeq
Throughout this study, we keep the width of the Gaussian kernel $w$ constant and equal to
1.2 pixels (corresponding to FWHM = 2.8 pixels).
Then we apply the same polynomial method to the 3$\times$3 patch around the brightest
pixel of the smoothed image. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star.

@@

Consider the case that we have a good estimate of the PSF at
the position of the faint star under consideration. Therefore, we 
can modify the smoothing step in the following way. Instead of smoothing
with a Gaussian kernel with a fixed size, we convolve the image of star
with the full postage stamp of the PSF. Then, we apply the same polynomial
fitting to the 3$\times$3 patch around the brightest pixel of the convolved image.

The procedure of convolving the image of star with the PSF results in a
more well-sampled, which eventually, leads to having more
information regarding the centroid of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple 2d polynomial will provides a better fit 
since convolution with the PSF makes the variation of the image across the 3 pixels
very smooth.

\item {\bf psf profile fitting}\quad

In addition to the methods mentioned so far, we examine fitting stars with the exact
 PSF profile. That is, in our Cramer-Rao bound saturation tests, That is, we find the best
 estimates of flux and centroid by
optimizing the $\chi^{2}$. We expect this method to perform better in determing the centroid
 of stars. In theory, it always saturates the CRLB.

@@

Moreover, it is more convenient to work with the signal-to-noise ratio
(hereafter denoted by SNR) instead of the variance of the Gaussian noise.
We define SNR as the ratio of the mean and variance of the distribution
which the flux estimator is drawn from. Assume that the total flux from
the point source is $F$, and that the PSF at the $i$-th pixel is given
by $P_{i}$. Therefore the brightness of the $i$-th pixel is drawn from
a Gaussian distribution $p(I_{i}) = \mathcal{N}(FP_{i},\sigma^{2})$.

The optimal estimator of flux (one which saturates Cramer-Rao bound),
is the weighted average $\tilde{F}=\sum_{i}I_{i}P_{i}$. Knowing the
distribution of $I_{i}$, it can be shown that 
\beq
p(\tilde{F}) = \mathcal{N}(F , \frac{\sigma^{2}}{\sum_{i}P_{i}^{2}}),
\eeq  
which leads us to
\beq
\begin{array}{l}
\text{SNR} $=$ \frac{\sqrt{\sum_{i} P_{i}^{2}}}{\sigma}.
\end{array}
\label{snr}
\eeq

Note that (\ref{snr}) is valid in the limit of sky-dominated regimes,
and that is the limit we are interested in throughout our investigation.
 In the case of our simulations, SNR given in (\ref{snr}) can be analytically
 expressed in terms of the per pixel uncertainty
$\sigma$, FWHM $\gamma$, and also $\beta$. We fix $\beta$ at the fiducial value of 2.5.
In this case we have
\beq
\text{SNR} = \frac{0.478}{\sigma \gamma}.
\label{snr2}
\eeq

@@

Given the PSF model that we have adopted for our simulations, we need 
to derive an expression for the CRLB as a function of the size, and SNR of stars.
 We know that the CRLB is given by the inverse of 
the Fisher information matrix $\mathcal{F}$. Thus, in order to find the CRLB,
 it is sufficient to compute the Fisher matrix.
Let us assume that we have $M$ observables $\{f_{1}, ... , f_{M}\}$, each
related to $B$ model parameters $\{\theta_{1} , ... , \theta_{B}\}$. Assuming
uncorrelated Gaussian error $\sigma_{m}$ for each observable $f_{m}$, elements
of the Fisher matrix $\mathcal{F}_{ij}$ are given by
\beq
\mathcal{F}_{ij} = \sum_{m}\frac{1}{\sigma_{m}^{2}}\frac{\partial f_{m}}{\partial \theta_{i}}\frac{\partial f_{m}}{\partial \theta_{j}}
\label{fisher}
\eeq

@@

In the second set, we generate 100,000 17$\times$17 postage-stamps
of Moffat profile, with values of $\gamma$ randomly distributed 
between 2 pixels and 6 pixels, and with centroids drawn randomly within 
the central pixel. We choose four values for SNR: 5, 10, 20, and 40. 
For each SNR, and for each postage-stamp with a given $\gamma$, 
Gaussian noise, with variance corresponding to SNR and $\gamma$ through equation 
(\ref{snr2}), is added to the each postage-stamp.

@@

In this experiment, after finding the centroiding error for each method,
we compute the RMSE in bins of SNR in order to compare it to the CRLB. 
Results of the first experiment are shown in Figures [\ref{1}], [\ref{2}],
[\ref{3}], [\ref{4}]. All methods deliver results
with RMSE larger for fainter stars.
In case of 3$\times$3 polynomial centroiding [\ref{1}], RMSE becomes
quite large as we move towards fainter stars in our simulation.
For stars with larger $\gamma$, centroid estimate from the naive polynomial
method show large deviation from the CRLB. By increasing $\gamma$,
RMSE deviates further away from the CRLB.

However, RMSE from the modified centroiding method I [\ref{2}],
is very close to the CRLB except at the very small values of SNR (SNR $<$ 10).
As we increase $\gamma$ from 2 pixels to 2.8 pixels, RMSE gets closer
to the CRLB. Besides, Also, when we increase $\gamma$ to
4 and 5.6 pixels, RMSE gets farther from the CRLB. For stars with $\gamma$ = 2 pixels, 
the rate at which the RMSE from this method drops
eventually becomes smaller than the constant rate at which the CRLB
decreases with increaing SNR. The reason for this is that even after smoothing
the with a Gaussian kernel, the image is still not well-sampled, and smooth enough
for a 2d polynomial fitting to provide highly accurate centroiding estimates.
For stars with $\gamma=2.8$ pixels, since the FWHM of the Gaussian kernel matches that 
of the underlying simulations, the smooth images are well-sampled and therefore, 
the method is able to deliver estimates very close to saturating the CRLB. As $\gamma$
gets higher for the simulations, the convolved images are not as well-sampled as those 
whose $\gamma$ matches the FWHM of the smoothing kernel and as a result, we loose
some information by fitting a polynomial to only a 3$\times$3 patch. That is the 
reason for why we observe slight deviation from the CRLB for stas with $\gamma$ = 4 pixels, and 
slightly more deviations as we increase $\gamma$ to 5.6 pixels.

As we expected, the RMSE from centroiding by fitting the exact PSF model [\ref{4}]
lies on the CRLB except for simulations with SNR = 5 where the RMSE gets slightly pulled away
from the CRLB due to presence of a few outliers. Figure [\ref{3}] demonstrates that even the modified centroiding
 method II is able deliver centroiding estimates as accurate as PSF profile fitting, with the exception
of simulated stars with $\gamma$ = 2 pixels. For stars with $\gamma$ = 2 pixels, the 3$\times$3 polynomial
fitting fails to saturate the CRLB since the PSF is undersampled. For simulated images higher $\gamma$, convolving the
data with the PSF result in images that are well-sampled around the brightest pixel. This allows the 
polynomial centroiding to deliver highly accurate results that can saturate the CRLB for simulated
data with SNR $\ga$ 10.

@@

In this experiment, after finding the centroiding error for each method, we
compute the RMSE in bins of $\gamma$ in order to compare it to the CRLB. 
Behavior of error as a function of $\gamma$ for different values of SNR,
is shown in figures [\ref{5}], [\ref{6}], and [\ref{7}], and [\ref{8}].

@@

The 3$\times$3 modified polynomial method I [\ref{6}], results in RMSE very close to the CRLB.
For all four values of SNR, as we increase $\gamma$ from 2 pixels to 3 pixels,
RMSE gets slightly closer to the CRLB since the method starts to perform
slightly better as we move away from undersampled stars, and as the FWHM of the smoothing kernel
gets closer to that of simulated images.
After approximately 3 pixels, increasing $\gamma$ result in deviation of RMSE of the method from the CRLB.
This is a characteristic of 3$\times$3 method as we apply it to a smoothed image which is still not sufficiently
well sampled.  Furthermore, increasing SNR from 5 to 40 makes the RMSE (as a function of $\gamma$) get
closer to the CRLB.

Again The RMSE from centroiding by fitting the exact PSF model as a function of FWHM
perfectly lies on the CRLB [\ref{8}] except at SNR = 5 where the RMSE slightly deviates
from the CRLB due to presence of a few outliers. Thus, centroid estimates from fitting
the exact PSF model always saturate the CRLB. Once again, we observe that the centroid
estimates found by the 3$\times$3 modified polynomial method II saturate the CRLB with the
exception of simulated stars with SNR = 5, or $\gamma$ very close to 2 pixels. [\ref{7}].

@@

A perfectly optimal stellar centroiding algorithm must saturate the fundamental Cramer-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroid of stars. In practice however,
that is only achievable when we have a very good estimate of the PSF. Since, we do not always 
know the exact PSF profile, we make use of various approximate centroiding algorithms. In this work, we
studied how different methods get close to saturating the CRLB for relatively low signal-to-noise ratio
stars.

We focused on examples from two classes of centroiding algorithms. The first class contains fast and approximate
methods that do not require any knowledge of the PSF at the poistion of stars. Of methods that belong to this class
we consider a simple method for centroiding stars based on fitting a low order
polynomial to a 3$\times$3 patch of star images, once smoothed by a Gaussian kernel of fixed width, and once without any smoothing step. 
The second class of centroiding algorithms contains methods which require knowledge of the PSF (or having a good estimate
of the PSF) at the position of stars. We considered two examples from this class. The first example
was the 3\times3 polynomial centroiding applied to the image of the star convolved with the PSF, and the
second example is the PSF profile fitting. In terms of saturating the Cramer-Rao bound we compared 
the performances of these mthods against each other. We call the 3$\times$3 polynomial method applied 
to the image convolved with a fixed Gaussian kernel (the PSF), 3$\times$3 polynomial method I (II).

%Furthermore, we compared the performance of these methods with the naive 3$\times$3 polynomial
%centroiding, 3$\times$3 quartic approximation, moment-based method, and PSF fitting.
%In order to make this comparison, we performed some tests on a large number of
%simulated data, and investigated to see if these estimators can saturate the Cramer-Rao
%lower bound, or how close to they get to saturating the bound. 
Our results suggest that in all ranges of FWHM and SNR, the PSF fitting method returns 
centroid estimates that saturate the CRLB; with the exception of SNR being less than 10, 
in which a few centroiding ouliers caused the RMSE to slightly, but not significantly, deviate
from the CRLB. We found that the estimates found by 3$\times$3 polynomial method are far away from
saturating the CRLB. On the other hand, the RMSE of centroid estimates of 
3$\times$3 modified polynomial method I are much closer to saturating the CRLB.
The performance of this two method however, is limited by
two important factors: \emph{(i)} signal-to-noise ratio, and \emph{(ii)} size of the stars.

The 3$\times$3 modified polynomial method I technique only takes advantage of the information contained in a 3$\times$3
patch around the brighest pixel of the smoothed image which is only well-sampled when the FWHM of the simulated star
 matches that of the smoothing kernel. Thus, when we apply 
these methods to find the centroid of stars with larger FWHM, a certain amount of
information (encoded in the Cramer-Rao lower bound) is lost, and therefore, the RMSE
of these methods deviates from the CRLB. This deviation gets even larger at lower
signal-to-noise ratios. Besides, the performance of this method slightly degrades
in the case of undersampled stars (with FWHM close two 2 pixels). 
Presence of noise is another limitting factor.
Although this method is able to get close to saturating the CRLB in a wide range of
signal-to-noise ratios, It is not reliable in the case of extremely faint stars
(SNR less than 10). This is mostly due to the fact that in the presence of noise, 
the brightest pixel of image, does not necessarily contain the centroid of stars even after smoothing 
the image.

Once we modify the polynomial method even further by smoothing the image with the full postage stamp
of the PSF, we obtain results that saturate the CRLB in a wide range of PSF sizes and noise levels.  
This is due to the fact that once the images are convolved with the right PSF, they become so well-sampled and smooth
that fitting a 2d polynomial to the 3$\times$3 patch around the brightest pixel of the smooth image can deliver
results as accurate as results obtained from fitting a PSF profile. The only exception where we observed deviaiotn from
saturating the CRLB occured in the case of extremely faint star (SNR $\la$ 10), and undersampled PSF (FWHM \sim 2 pixels).
One advantage of this method over PSF fitting method is that it is fast, and that it is able to saturate the CRLB in 
a wide range of conditions. Although this method is more accurate than 3$\times$3 modified polynomial I, but it is
slightly slower because it requires convolution of the image with the full potage stamp of the PSF.

In this investigation we showed that PSF fitting is always better at centroiding stars.
However, this method has its own disadvantegouses since we do not always know the the exact PSF, and 
finding the centroid by PSF fitting is computationally expensive whereas by employing the 3$\times$3 modified polynomial I
technique into large scale astronomical surveys can considerably reduce the computational cost
of initial astrometry of these surveys. This method can roughly be thousand
times faster than centroiding by model fitting. Even in the case that we have a good estimate of the PSF, the 
3$\times$3 modified polynomial II is a much faster algorithm.

Moreover, it is important to note that the PSF fitting method can be made faster by only keeping the 
term proportional to the product of the PSF model and the image in $\chi^2$. This is because 
the other terms are not sensitive to the position of centroid. However, this only allows
us to vary the poition of centroid, and not the flux, while fitting the PSF model
to the star.

Modifying the $\chi^2$ such that it contains the product of the PSF and image, 
provides a nice interpretation for the PSF fitting method. It corresponds to oversampling
of the image with the correct PSF, and finding the centroid in the oversampled image afterwards.
On the other hand, one can think of the initial smoothing step in 3$\times$3 methods
such as modified polynomial and quartic approximation methods, as oversampling of the 
image with an approximate Gaussian PSF. The fitting and approximate methods are different 
in that after upsamlping, in approximate methods, our knowledge is limited to the 3$\times$3 
patch around the brightest pixel and as a result, we loose some information.

Having a reasonable PSF model should always deliver better centroid estimates, but
over a certain range of low signal-to-noise ratios, one can achieve sensibly accurate
results by employing this simple 3$\times$3 method without making any assumption about
the PSF model.

In this investigation we narrowed our focus on a set of data simulated from a very specific
model. Although there are various cases where Moffat profiles provide reasonable
representation of point spread function, these models are not generic enough to let us
reach a more general conclusion. Another important, and untapped, area of study
would be devising a model that infers the centroid of stars and point spread function
(in its full extent, not at the catalog level) across an astronomical image simultaneously.
This is beyond the scope of this study.

@@

\begin{figure}[!htb]
\minipage{.8\textwidth}
  \includegraphics[width=\linewidth]{snr_psfpix28poly.png}
\endminipage
\caption{Scatter plots showing the relation between error in centroid measurement from the  and the signal-to-noise ratio of stars, with FWHM of : 2 (upper left), 2.8 (upper right), 4 (lower left), and 5.6 (lower right) pixels. In each scatter plot, the blue solid line represents the root-mean-squared-error, and the green dashed line represents CRLB.}\label{2}
\end{figure}

>>>>

In this paper, we study how optimal various techniques are at centroiding 
faint stars. That is, we apply a number of centroiding methods 
to a large number of simulated faint stars with different signal-to-noise ratio and
 size realizations. Error from star centroiding methods, will always have 
a theoretically-set lower bound, known as the Cramer-Rao lower bound
(hereafter denoted by CRLB) which has an inverse relation with the
signal-to-noise-ratio of stars.

Given the analytic expression for the PSF model,
we derive an expression for the fundamental lower bound on the centroiding error as
a function of the parameters of the PSF model (\eg size),
and signal-to-noise-ratio of stars. We create two sets of simulations for which we 
know the CRLB, one with variable (constant) signal-to-noise ratio (FWHM), and one 
with variable (constant) FWHM (signal-to-noise ratio). After applying
different centroiding methods to the simulations,
we investigate how close these methods can get to saturating the CRLB for
various ranges of background noise level and PSF FWHM.

In this work, we focus on four centroiding methods. The first method
estimates the centroid of a star by a fitting a 2d polynomial to the 3$\times$3 patch around 
the brightest pixel of the image. The second method centroids stars by
 smoothing the image of stars by a Gaussian kernel of a fixed size,
 and then applying the same 3$\times$3 polynomial trick to the smooth
 image. This method is fast and does not require any knowledge of the 
PSF. In the third method, the 3$\times$3 polynomial centroiding is applied to the
 image of star convolved with the full PSF postage stamp. The last method we consider,
 is fitting a PSF profile, assuming that we have a good PSF estimate, to the star.
%We compare the accuracy of this method with, the naive 3$\times$3 second-order 
%polynomial method, the 3$\times$3 quartic approximation
%method used in SDSS photometric pipeline, and fitting a PSF model to the 
%star.

This paper is structured as follows. In section \ref{sec:method},
we give a brief overview of the centroiding methods used in our investigation.
In section \ref{sec:data}, we discuss the simulated data, and the Cramer-Rao bound
 on centroiding error. In section \ref{sec:results}, we compare the performances of the methods
discussed in \ref{sec:method}, with the CRLB derived in \ref{sec:data}. Finally,
 we discuss and conclude in \ref{sec:discussion}.

\section{Methods}\label{sec:Centroiding methods}

@@

Now, we modify the simple polynomial method explained above,
by smoothing the image of the star with a Gaussian kernel
\beq
k(\mathbf{x}) = \frac{1}{2\pi w^2}\exp(-\mathbf{x}^{2}/2w^{2}).
\eeq
Throughout this study, the full-width at half-maximum of the Gaussian kernel is held at
a fixed value of 2.8 pixels (corresponding to w $\simeq$ 1.2 pixels).

Then we apply the same polynomial method to the 3$\times$3 patch around the brightest
pixel of the smoothed image. Therefore, for a given star and a smoothing kernel,
the outcome of equation (\ref{linearfit}) can be
plugged into equation (\ref{center}) to find the centroid estimate
of the star.

@@

Now let us consider the case in which we have a good estimate of the PSF at
the position of the faint star under consideration. Therefore, we 
can modify the smoothing step in the following way. Instead of smoothing
with a Gaussian kernel of a fixed size, we convolve the image of star
with the full postage stamp of the PSF. Then, we apply the same polynomial
fitting to the 3$\times$3 patch around the brightest pixel of the convolved image.

The procedure of convolving the image of star with the PSF results in a
more well-sampled image, which eventually, leads to having more
information regarding the centroid of stars in the 3$\times$3 patch around the
brightest pixel. In addition, a simple 2d polynomial will provide a better fit 
since convolution with the PSF makes the variation of the brightness of the image
 across the 3$\times$3 patch very smooth.

\item {\bf PSF profile fitting}\quad

In addition to the methods mentioned so far, we examine fitting an exact
 PSF profile to the stars. That is, in our Cramer-Rao bound saturation tests, we find the best
 estimates of flux and centroid by optimizing the $\chi^{2}$.
 We expect this method to perform better in determing the centroid
 of stars. In theory, this method should always saturate the CRLB.

@@

Moreover, it is more convenient to work with the signal-to-noise ratio
(hereafter denoted by SNR) instead of the variance of the Gaussian noise.
We use the definition of SNR according to which, SNR is given by the ratio
 of the mean and variance of the distribution
which the flux estimator is drawn from. Assume that the total flux from
the point source is $F$, and that the PSF at the $i$-th pixel is given
by $P_{i}$. Therefore the brightness of the $i$-th pixel is drawn from
a Gaussian distribution $p(I_{i}) = \mathcal{N}(FP_{i},\sigma^{2})$.

The optimal estimator of flux (one that saturates Cramer-Rao bound),
is the weighted average $\tilde{F}=\sum_{i}I_{i}P_{i}$. Knowing the
distribution of $I_{i}$, it can be shown that 
\beq
p(\tilde{F}) = \mathcal{N}(F , \frac{\sigma^{2}}{\sum_{i}P_{i}^{2}}),
\eeq  
which leads us to
\beq
\begin{array}{l}
\text{SNR} $=$ \frac{\sqrt{\sum_{i} P_{i}^{2}}}{\sigma}.
\end{array}
\label{snr}
\eeq

Note that (\ref{snr}) is valid in the limit of sky-dominated regimes,
and that is the limit we are interested in throughout our investigation.
 In the case of our simulations, SNR given in (\ref{snr}) can be analytically
 expressed in terms of the per pixel uncertainty
$\sigma$, FWHM $\gamma$, and also $\beta$. We hold $\beta$ fixed at the fiducial value of 2.5.
In this case we have
\beq
\text{SNR} = \frac{0.478}{\sigma \gamma}.
\label{snr2}
\eeq

@@

Given the PSF model that we have adopted for our simulations, we can derive
 an expression for the CRLB as a function of the size, and SNR of stars.
 We know that the CRLB is given by the inverse of 
the Fisher information matrix $\mathcal{F}$. Thus, in order to find the CRLB,
 it is sufficient to compute the Fisher matrix.
Let us assume that we have $M$ observables $\{f_{1}, ... , f_{M}\}$, each
related to $B$ model parameters $\{\theta_{1} , ... , \theta_{B}\}$. Assuming
uncorrelated Gaussian error $\sigma_{m}$ for each observable $f_{m}$, elements
of the Fisher matrix $\mathcal{F}_{ij}$ are given by
\beq
\mathcal{F}_{ij} = \sum_{m}\frac{1}{\sigma_{m}^{2}}\frac{\partial f_{m}}{\partial \theta_{i}}\frac{\partial f_{m}}{\partial \theta_{j}}
\label{fisher}
\eeq

@@

In the second set, we generate 100,000 17$\times$17 postage-stamps
of Moffat profile, with values of $\gamma$ uniformly distributed 
between 2 and 6 pixels, and with centroids drawn randomly within 
the central pixel. We choose four values for SNR: 5, 10, 20, and 40. 
For each SNR, and for each postage-stamp with a given $\gamma$, 
Gaussian noise, with variance corresponding to SNR and $\gamma$ through equation 
(\ref{snr2}), is added to each postage-stamp.

@@

In this experiment, after finding the centroiding error for each method,
we compute the RMSE in bins of SNR in order to compare it to the CRLB. 
Results of the first experiment are shown in Figures [\ref{1}], [\ref{2}],
[\ref{3}], and [\ref{4}]. All methods deliver results
with RMSE larger for fainter stars.
In case of 3$\times$3 polynomial centroiding [\ref{1}], RMSE becomes
quite large as we move towards fainter stars in our simulation.
For stars with larger $\gamma$, centroid estimates from the naive 3$\times$3
 polynomial method show large deviation from the CRLB. By increasing $\gamma$,
RMSE deviates further away from the CRLB.

However, RMSE from the modified centroiding method I [\ref{2}],
is very close to the CRLB except at the very small values of SNR (SNR $<$ 10).
As we increase $\gamma$ from 2 pixels to 2.8 pixels, RMSE gets closer
to the CRLB. When we increase $\gamma$ to
4 and 5.6 pixels, RMSE gets farther from the CRLB. For stars with $\gamma$ = 2 pixels, 
the rate at which the RMSE from this method drops
eventually becomes smaller than the constant rate at which the CRLB
decreases with increaing SNR. The reason for this is that even after smoothing
the with a Gaussian kernel, the image is still relatively undersampled, and not smooth enough
for a 2d polynomial fitting to provide highly accurate centroiding estimates.
For stars with $\gamma=2.8$ pixels, since the FWHM of the Gaussian kernel matches that 
of the PSF of underlying simulations, the smooth images are well-sampled and therefore, 
the method is able to deliver estimates extremely close to saturating the CRLB. As $\gamma$
gets higher for the simulations, the convolved images are not as well-sampled as those 
whose $\gamma$ matches the FWHM of the smoothing kernel and as a result, we loose
some information by fitting a polynomial to a 3$\times$3 patch. Therefore,
 we observe slight deviation from the CRLB for stars with $\gamma$ = 4 pixels, and 
slightly more deviation as we increase $\gamma$ to 5.6 pixels.

As we expected, the RMSE from centroiding by fitting the exact PSF model [\ref{4}]
lies on the CRLB except for simulations with SNR = 5 where the RMSE gets slightly pulled away
from the CRLB due to presence of a few outliers. Figure [\ref{3}] demonstrates that even the modified centroiding
 method II is able deliver centroiding estimates as accurate as PSF profile fitting, with the exception
of simulated stars with $\gamma$ = 2 pixels. For stars with $\gamma$ = 2 pixels, although the 3$\times$3 polynomial
fitting gets close to saturating the CRLB, it fails to saturate the CRLB since the PSF is undersampled. For simulated images with higher $\gamma$, convolving the
data with the PSF result in images that are well-sampled around the brightest pixel. This allows the 
polynomial centroiding to deliver highly accurate results that can saturate the CRLB for simulated
data with SNR $\ga$ 10.

@@

In this experiment, after finding the centroiding error for each method, we
compute the RMSE in bins of $\gamma$ in order to compare it to the CRLB. 
Behavior of error as a function of $\gamma$ for different values of SNR,
is shown in figures [\ref{5}], [\ref{6}], [\ref{7}], and [\ref{8}].

@@

The 3$\times$3 modified polynomial method I [\ref{6}], results in RMSE very close to the CRLB.
For all four values of SNR, as we increase $\gamma$ from 2 pixels to 3 pixels,
RMSE gets slightly closer to the CRLB since the method starts to perform
slightly better as we move away from undersampled stars, and as the FWHM of the smoothing kernel
gets closer to that of the simulated images.
After approximately 3 pixels, increasing $\gamma$ result in deviation of RMSE of the method from the CRLB.
This is a characteristic of 3$\times$3 method as we apply it to a smoothed image which is still not sufficiently
well-sampled.  Furthermore, increasing SNR from 5 to 40 makes the RMSE (as a function of $\gamma$) become
closer to the CRLB.

Once again, the RMSE from centroiding by fitting the exact PSF model as a function of FWHM
perfectly lies on the CRLB except at SNR = 5 where the RMSE slightly deviates
from the CRLB due to presence of a few outliers [\ref{8}].
. Thus, centroid estimates from fitting
the exact PSF model always saturate the CRLB. Once again, we observe that the centroid
estimates found by the 3$\times$3 modified polynomial method II saturate the CRLB with the
exception of simulated stars with SNR = 5, or $\gamma$ very close to 2 pixels [\ref{7}].

@@

An optimal stellar centroiding algorithm must saturate the fundamental Cramer-Rao lower 
bound. That is, in all ranges of background noise level, size, radial light profile,
and shape, it must preserve information about the centroid of stars. In practice however,
this is only achievable when we have a reasonably good estimate of the PSF. Since, we do not always 
know the exact PSF profile, we make use of various approximate centroiding algorithms. In this work, we
studied how different methods get close to saturating the CRLB for relatively low signal-to-noise ratio
stars.

We focused on examples from two classes of centroiding algorithms. The first class contains fast and approximate
methods that do not require any knowledge of the PSF at the position of stars. Of methods that belong to this class,
we consider a simple method for centroiding stars based on fitting a low order
polynomial to a 3$\times$3 patch of star images, once smoothed by a Gaussian kernel of fixed width, and once without any smoothing step. 
The second class of centroiding algorithms contains methods that require knowledge of the PSF (or having a good estimate
of the PSF) at the position of stars. We considered two examples from this class. The first example
was the 3$\times$3 polynomial centroiding applied to the image of the star convolved with the PSF, and the
second example is the PSF profile fitting. In terms of saturating the Cramer-Rao bound, we compared 
the performances of these methods against each other. Throughout this work, we called the 3$\times$3 polynomial method applied 
to the image convolved with a fixed Gaussian kernel ( convolved with the PSF), 3$\times$3 modified polynomial method I (II).

%Furthermore, we compared the performance of these methods with the naive 3$\times$3 polynomial
%centroiding, 3$\times$3 quartic approximation, moment-based method, and PSF fitting.
%In order to make this comparison, we performed some tests on a large number of
%simulated data, and investigated to see if these estimators can saturate the Cramer-Rao
%lower bound, or how close to they get to saturating the bound. 
Our results suggest that in all ranges of FWHM and SNR, the PSF fitting method returns 
centroid estimates that saturate the CRLB; with the exception of SNR being less than 10, 
in which a few centroiding ouliers caused the RMSE to slightly, but not significantly, deviate
from the CRLB. We found that the estimates found by 3$\times$3 polynomial method are far away from
saturating the CRLB. On the other hand, the RMSE of centroid estimates of 
3$\times$3 modified polynomial method I are much closer to saturating the CRLB.
The performance of this method however, is limited by
two important factors: \emph{(i)} signal-to-noise ratio, and \emph{(ii)} PSF size.

The 3$\times$3 modified polynomial method I technique only takes advantage of the information
 contained in a 3$\times$3 patch around the brighest pixel of the smoothed image which is only
 well-sampled when the FWHM of the simulated image of star
 matches that of the smoothing kernel. Thus, when we apply 
this method to find the centroid of stars with larger FWHM, certain amount of
information (encoded in the Cramer-Rao lower bound) is lost, and therefore, the RMSE
of these methods deviates from the CRLB. This deviation becomes even larger at lower
signal-to-noise ratios. Besides, the performance of this method slightly degrades
in the case of undersampled stars (with FWHM close two 2 pixels). 
Presence of noise is another limitting factor.
Although this method is able to get close to saturating the CRLB in a wide range of
signal-to-noise ratios, It is not reliable in the case of extremely faint stars.
 This is partly due to the fact that in the presence of noise, 
the brightest pixel of image does not necessarily contain the centroid of stars even after smoothing 
the image.

Once we modify the polynomial method even further by smoothing the image with the full postage stamp
of the PSF, we obtain results that saturate the CRLB in a wide range of PSF sizes and noise levels.  
This is due to the fact that once the images are convolved with the right PSF, they become so well-sampled and smooth
that fitting a 2d polynomial to the 3$\times$3 patch around the brightest pixel of the smooth image is sufficient for us to obtain
results as accurate as those from fitting a PSF profile. The only conditions in which deviation from
saturating the CRLB occures are very low signal-to-noise ratios (SNR $\la$ 10), and undersampled PSF (FWHM $\sim$ 2 pixels).
One advantage of this method over PSF fitting method is that it is fast, and that it is able to saturate the CRLB in 
a wide range of conditions. In the case that we have a good estimate of the PSF, the 
3$\times$3 modified polynomial II is a much faster algorithm. 
Although this method is more accurate than 3$\times$3 modified polynomial I, it is
slightly slower because it requires convolution of the image with the full potage stamp of the PSF.

In this investigation we showed that PSF fitting is always better at centroiding stars.
However, this method has its own disadvantages. First, we do not always know the the exact PSF. Second, finding
 the centroid by profile fitting is computationally expensive, whereas by employing any of the 3$\times$3 modified polynomial 
techniques into large scale astronomical surveys can considerably reduce the computational cost
of initial astrometry of the point sources detected by these surveys.

Moreover, it is important to note that the PSF fitting method can be made faster by only keeping the 
term proportional to the dot product of the PSF model and the image in $\chi^2$. This is because 
the other terms are not sensitive to the position of the centroid. However, this only allows
us to vary the poition of centroid, and not the flux, while fitting the PSF model
to the star.

Modifying the $\chi^2$ such that it contains the dot product of the PSF and image, 
provides a nice interpretation for the PSF fitting method. It corresponds to searching for the 
centroid in a very high resolution grid which the dot product of the image and the PSF model is
sensitive to. On the other hand, one can think of the initial smoothing step in the first modified 3$\times$3 method 
as upsampling of the image around the center through convolving it with an approximate Gaussian PSF. 
The PSF profile fitting and the first modified method are different in that when there is mismatch between the
widths of the smoothing kernel and the PSF, we loose some information by employing a 3$\times$3 polynomial trick.
When we have the advantage of knowing the PSF, this isuue can be resolved by employing the second modified polynomial method.

Having a reasonable PSF model should always deliver better centroid estimates, but
over a certain range of low signal-to-noise ratios and PSF sizes, one can achieve sensibly
 accurate results by employing a simple 3$\times$3 method after smooting the image with
 a Gaussian kernel of fixed width, and without making any assumption about the PSF model.

In this investigation we narrowed our focus on a set of data simulated from a particular
PSF profile. Although there are various cases where Moffat profiles provide reasonable
representation of point spread function, these profiles are not generic enough to let us
reach a more general conclusion. Another important, and untapped, area of study
would be devising a model that infers the centroid of stars and point spread function
(in its full extent, not at the catalog level) across an astronomical image simultaneously.
This is beyond the scope of this study.

@@

\begin{figure}[!htb]
\minipage{.8\textwidth}
  \includegraphics[width=\linewidth]{snr_psfpix28poly.png}
\endminipage
\caption{Scatter plots showing the relation between error in centroid measurement from the 3$\times$3 modified polynomial method and the signal-to-noise ratio of stars, with FWHM of : 2 (upper left), 2.8 (upper right), 4 (lower left), and 5.6 (lower right) pixels. In each scatter plot, the blue solid line represents the root-mean-squared-error, and the green dashed line represents CRLB.}\label{2}
\end{figure}

\begin{figure}[!htb]
\minipage{.8\textwidth}
  \includegraphics[width=\linewidth]{snr_psfpoly.png}
\endminipage
\caption{Scatter plots showing the relation between error in centroid
measurement from the 3$\times$3 modified polynomial method II and the signal-to-noise
ratio of stars, with FWHM of : 2 (upper left), 2.8 (upper right), 4 (lower left),
and 5.6 (lower right) pixels. In each scatter plot, the blue solid line represents the root-mean-squared-error, and the green dashed line represents CRLB.}\label{3}
\end{figure}
